### 2019-12-05 writing/paper/paper.tex
Start paper

\documentclass[twocolumn]{article}

\title{WorldGAN: 3D Environment Generation Using Deep Adversarial Networks}

\author{Tim Whitaker}

\begin{document}

	\maketitle

	\section{Abstract}

	\section{Introduction}

	\subsection{Objectives}

	\section{Background}

	\section{Methodology}

	Building the UI app.

	Building the training set.

	First round of training for pix2pix and cyclegan.

	Augmenting the training set.

	Second round of training for pix2pix and cyclegan.

	Building the test set.

	Visualizations of results with 3d software.

	\section{Results}

	\section{Conclusion}

\end{document}